---
title: "Liver Disease Analysis"
author: "Sean Johnson"
date: "2021-11-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library
```{r, warning=FALSE,message=FALSE}
library(dplyr)
```

# Data
```{r, message=FALSE}
#data was downloaded from the IC repository
hcv = readr::read_csv(here::here("data","hcvdat0.csv"))
```

# Quick Overview
```{r}
summary(hcv)
```

# Data Wrangling

## Removing NAs
```{r}
df = as.data.frame(na.omit(hcv))
1 - nrow(df)/nrow(hcv)
```

## Renaming
```{r}
#cleaning
df<-df%>%
  mutate(Category = case_when(Category == "0=Blood Donor" ~0,
                              Category == "0s=suspect Blood Donor" ~1,
                              Category == "1=Hepatitis" ~2,
                              Category == "2=Fibrosis" ~3,
                              Category == "3=Cirrhosis" ~4),
         Sex = case_when(Sex=="m" ~0,
                         Sex=="f" ~1))
```

## Drop ID Col
```{r}
df <-df%>%
  select(-...1)
```

## Formatting
```{r}
df[,1] = as.numeric(df[,1])
df[,3] = as.numeric(df[,3])
```

```{r}
ind0 = (df$Category==0)
ind1 = (df$Category==1)
ind = ind0 | ind1
self = df[ind,]
bad = df[ind1,]
allBad = df[!ind0,]
#checks
nrow(allBad)/nrow(df)
nrow(bad)/nrow(df)
```

# Subset of DF
```{r}
x = df
x<-x%>%
  mutate(Category = case_when(Category == 2 ~1,
                              Category == 3 ~1,
                              Category == 4 ~1,
                              TRUE ~0))%>%
  mutate(Category = as.factor(Category))

```

# Table for Predictive Modeling
```{r}
# table
table_accuracy = matrix(nrow=6,ncol=3)
colnames(table_accuracy) = c('Accuracy','Precision','Recall')
rownames(table_accuracy) = c('DTree','NB','SVM-Linerar','SVM-Polynomial','ANN','KNN')
table_accuracy
```
# K-fold Cross-Validation
```{r}
n = nrow(x)
k = 10
tail = n%/%k

set.seed(2)

rnd = runif(n)
rank = rank(rnd)
blk = (rank-1)%/%tail+1
blk = as.factor(blk)

print(summary(blk))
```
# Predictive Models

## Decision Tree
```{r}
set.seed(2)

all.acc = numeric(0)
all.pre = numeric(0)
all.rec = numeric(0)
for(i in 1:k){
  tree = rpart::rpart(Category~.,x[blk != i,],method="class")
  pred = predict(tree,x[blk==i,],type="class")
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2])/sum(confMat)
  pre = (confMat[1,1])/sum(confMat[1,])
  rec = (confMat[1,1])/sum(confMat[,1])
  all.acc = rbind(all.acc,acc)
  all.pre = rbind(all.pre,pre)
  all.rec = rbind(all.rec,rec)
}
j=1
print(mean(all.acc))
print(mean(all.pre))
print(mean(all.rec))
table_accuracy[j,1] = mean(all.acc)
table_accuracy[j,2] = mean(all.pre)
table_accuracy[j,3] = mean(all.rec)
```
## Naive Bayes
```{r}
set.seed(2)

all.acc = numeric(0)
all.pre = numeric(0)
all.rec = numeric(0)
for(i in 1:k){
  model = e1071::naiveBayes(Category~.,x[blk != i,],method="class")
  pred = predict(model,x[blk==i,],type="class")
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2])/sum(confMat)
  pre = (confMat[1,1])/sum(confMat[1,])
  rec = (confMat[1,1])/sum(confMat[,1])
  all.acc = rbind(all.acc,acc)
  all.pre = rbind(all.pre,pre)
  all.rec = rbind(all.rec,rec)
}
j=2
print(mean(all.acc))
print(mean(all.pre))
print(mean(all.rec))
table_accuracy[j,1] = mean(all.acc)
table_accuracy[j,2] = mean(all.pre)
table_accuracy[j,3] = mean(all.rec)
```
## SVM Linear
```{r}
set.seed(2)

all.acc = numeric(0)
all.pre = numeric(0)
all.rec = numeric(0)
for(i in 1:k){
  model = e1071::svm(Category~.,x[blk != i,],kernel="linear",type="C")
  pred = predict(model,x[blk==i,],type="class")
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2])/sum(confMat)
  pre = (confMat[1,1])/sum(confMat[1,])
  rec = (confMat[1,1])/sum(confMat[,1])
  all.acc = rbind(all.acc,acc)
  all.pre = rbind(all.pre,pre)
  all.rec = rbind(all.rec,rec)
}
j=3
print(mean(all.acc))
print(mean(all.pre))
print(mean(all.rec))
table_accuracy[j,1] = mean(all.acc)
table_accuracy[j,2] = mean(all.pre)
table_accuracy[j,3] = mean(all.rec)
```

## SVM Polynomial
```{r}
set.seed(2)

all.acc = numeric(0)
all.pre = numeric(0)
all.rec = numeric(0)
for(i in 1:k){
  model = e1071::svm(Category~.,x[blk != i,],kernel="polynomial",type="C")
  pred = predict(model,x[blk==i,],type="class")
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2])/sum(confMat)
  pre = (confMat[1,1])/sum(confMat[1,])
  rec = (confMat[1,1])/sum(confMat[,1])
  all.acc = rbind(all.acc,acc)
  all.pre = rbind(all.pre,pre)
  all.rec = rbind(all.rec,rec)
}
j=4
print(mean(all.acc))
print(mean(all.pre))
print(mean(all.rec))
table_accuracy[j,1] = mean(all.acc)
table_accuracy[j,2] = mean(all.pre)
table_accuracy[j,3] = mean(all.rec)
```
## Nerual Net
```{r}
set.seed(2)

all.acc = numeric(0)
all.pre = numeric(0)
all.rec = numeric(0)
for(i in 1:k){
  model = nnet::nnet(Category~.,x[blk != i,], size = 7, trace=FALSE, wgts=.1)
  pred = as.integer(predict(model, x[blk==i,]))
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1])/sum(confMat)
  pre = (confMat[1,1])/sum(confMat[1,])
  rec = (confMat[1,1])/sum(confMat[,1])
  all.acc = rbind(all.acc,acc)
  all.pre = rbind(all.pre,pre)
  all.rec = rbind(all.rec,rec)
}
j=5
print(mean(all.acc))
print(mean(all.pre))
print(mean(all.rec))
table_accuracy[j,1] = mean(all.acc)
table_accuracy[j,2] = mean(all.pre)
table_accuracy[j,3] = mean(all.rec)
```

## K-Nearest Neighbors
```{r}
set.seed(2)
n=5

trControl = caret::trainControl(method="cv",number=n)
x1 = x[,]
x1$Category = as.factor(x1$Category)
model = caret::train(Category ~ ., method = "knn", tuneGrid = expand.grid(k = 1:10), trControl = trControl, data = x1)
model
```


```{r}
set.seed(2)

aall.acc = numeric(0)
all.pre = numeric(0)
all.rec = numeric(0)
for(i in 1:k){
  tr = x1[blk != i,]
  te = x1[blk == i,]
  pred = class::knn(train = tr, test = te, cl = tr[,1], k=8)
  confMat = table(pred,x1$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2])/sum(confMat)
  pre = (confMat[1,1])/sum(confMat[1,])
  rec = (confMat[1,1])/sum(confMat[,1])
  all.acc = rbind(all.acc,acc)
  all.pre = rbind(all.pre,pre)
  all.rec = rbind(all.rec,rec)
}
j=6
print(mean(all.acc))
print(mean(all.pre))
print(mean(all.rec))
table_accuracy[j,1] = mean(all.acc)
table_accuracy[j,2] = mean(all.pre)
table_accuracy[j,3] = mean(all.rec)
```
# Table of Models
```{r}
tab <- round(table_accuracy,4)
tab <- as_tibble(tab)
tab$`Model Name` <-  c('DTree','NB','SVM-Linerar','SVM-Polynomial','ANN','KNN')
tab%>%
  select(`Model Name`,Accuracy,Precision,Recall)%>%
  Thematic::tabGT(table_title = "Model Performance")
```
## Write Out Info for further use
```{r}
write.table(tab, file = 'data/accuracy.txt', sep =' ', row.names = TRUE, col.names = TRUE)
write.table(x, file = 'data/main_df.txt', sep =' ', row.names = TRUE, col.names = TRUE)
```

# ~~~~~~~~~
 
# Analysis Liver Disease Stages

```{r}
x = allBad
```

# Table 
```{r}
table_accuracy = matrix(nrow=6,ncol=1)
colnames(table_accuracy) = c('Accuracy')
rownames(table_accuracy) = c('DTree','NB','SVM-Linerar','SVM-Polynomial','ANN','KNN')
table_accuracy
```

# K-Fold CV
```{r}
n = nrow(x)
k = 5
tail = n%/%k

set.seed(2)

rnd = runif(n)
rank = rank(rnd)
blk = (rank-1)%/%tail+1
blk = as.factor(blk)

print(summary(blk))
```
```{r}
#cannot have there be a categoryof 1 in holdout.
print(x$Category[blk==6])
```

# Predictive Models
## Decision Tree
```{r}
set.seed(2)
all.acc = numeric(0)
for(i in 1:k){
  tree = rpart::rpart(Category~.,x[blk != i,],method="class")
  pred = predict(tree,x[blk==i,],type="class")
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2]+confMat[3,3]+confMat[4,4])/sum(confMat)
  all.acc = rbind(all.acc,acc)
}
j=1
print(mean(all.acc))
table_accuracy[j,1] = mean(all.acc)
```
 
## Naive Bayes
```{r}
set.seed(2)
all.acc = numeric(0)
for(i in 1:k){
  model = e1071::naiveBayes(Category~.,x[blk != i,],method="class")
  pred = predict(model,x[blk==i,],type="class")
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2]+confMat[3,3]+confMat[4,4])/sum(confMat)
  all.acc = rbind(all.acc,acc)
}
j=2
print(mean(all.acc))
table_accuracy[j,1] = mean(all.acc)

```

## SVM Linear
```{r}
set.seed(2)
all.acc = numeric(0)
for(i in 1:k){
  model = e1071::svm(Category~.,x[blk != i,],kernel="linear",type="C")
  pred = predict(model,x[blk==i,],type="class")
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2]+confMat[3,3]+confMat[4,4])/sum(confMat)
  all.acc = rbind(all.acc,acc)
}
j=3
print(mean(all.acc))
table_accuracy[j,1] = mean(all.acc)



```

## SVM Polynomial
```{r}
set.seed(2)
all.acc = numeric(0)
for(i in 1:k){
  model = e1071::svm(Category~.,x[blk != i,],kernel="polynomial",type="C")
  pred = predict(model,x[blk==i,],type="class")
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2]+confMat[3,3]+confMat[4,4])/sum(confMat)
  all.acc = rbind(all.acc,acc)
}
j=4
print(mean(all.acc))
table_accuracy[j,1] = mean(all.acc)
```

## Neural Network
```{r}
set.seed(2)
all.acc = numeric(0)
for(i in 1:k){
  model = nnet::nnet(Category~.,x[blk != i,], size = 7, trace=FALSE, wgts=.1)
  pred = as.integer(predict(model, x[blk==i,]))
  confMat = table(pred,x$Category[blk==i])
  acc = (confMat[1,1])/sum(confMat)
  all.acc = rbind(all.acc,acc)
}
j=5
print(mean(all.acc))
table_accuracy[j,1] = mean(all.acc)
```

## K Nearest Neighbors
```{r}
set.seed(2)
n=5
trControl = caret::trainControl(method="cv",number=n)
x1 = x[,]
x1$Category = as.factor(x1$Category)
model = train(Category ~ ., method = "knn", tuneGrid = expand.grid(k = 1:10), trControl = trControl, data = x1)
model
```


```{r}
set.seed(2)
aall.acc = numeric(0)
for(i in 1:k){
  tr = x1[blk != i,]
  te = x1[blk == i,]
  pred = class::knn(train = tr, test = te, cl = tr[,1], k=10)
  confMat = table(pred,x1$Category[blk==i])
  acc = (confMat[1,1]+confMat[2,2]+confMat[3,3]+confMat[4,4])/sum(confMat)
  all.acc = rbind(all.acc,acc)
}
j=6
print(mean(all.acc))
table_accuracy[j,1] = mean(all.acc)
```

# Table
```{r}
tab = round(table_accuracy,4)
tab
```

# Write Out
```{r}
write.table(tab, file = 'data/accuracy_allBad.txt', sep =' ', row.names = TRUE, col.names = TRUE)
write.table(x, file = 'data/allBad_df.txt', sep =' ', row.names = TRUE, col.names = TRUE)
```

